---
title: "Multiple history workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Multiple history workflow}
  %\usepackage[UTF-8]{rtrackr}
---

```{r echo=F, results="hide"}
suppressPackageStartupMessages({
  library(kableExtra)
  library(rtrackr)
  library(networkD3)
})
```

### Setting up
Below is an example of a more complicated workflow, starting from multiple input files, to be used as an example of a full `rtrackr` workflow.

We will define and log simple dataframes for this example. For more information, please see [getting started](https://hamishgibbs.github.io/rtrackr/articles/getting_started.html) and [summarising data](https://hamishgibbs.github.io/rtrackr/articles/summarising_data.html).

```{r}
trackr_dir <- '~/Documents/Personal/trackr_dev/trackr_dir'

df1 <- data.frame(a = c('a', 'b', 'c'), b = c(1, 2, 3))
df2 <- data.frame(a = c('a', 'b', 'c'), b = c(4, 5, 6))
```
<details><summary>Details 1</summary>
**df1**
```{r, echo = F}
kable(df1) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
<details><summary>Details 2</summary>
**df2**
```{r, echo = F}
kable(df2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>

```{r}
df1 <- trackr_new(df1, trackr_dir = trackr_dir, timepoint_message = 'Start', suppress_success = TRUE)
df2 <- trackr_new(df2, trackr_dir = trackr_dir, timepoint_message = 'Start', suppress_success = TRUE)
```
<details><summary>Details 1</summary>
**df1**
```{r, echo = F}
kable(df1) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
<details><summary>Details 2</summary>
**df2**
```{r, echo = F}
kable(df2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>

### Workflow

```{r}
df <- rbind(df1, df2)

df <- df %>% dplyr::group_by(a) %>% 
  trackr_summarise(n = dplyr::n())
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
```{r}
df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'processing 1', suppress_success = TRUE)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
```{r}
df <- df %>% 
  dplyr::mutate(n = n + 20)

df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'processing 1', suppress_success = TRUE)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>

### Results
Inspect the full processing chain of a single record with [`trackr_lineage()`](https://hamishgibbs.github.io/rtrackr/reference/trackr_lineage.html) and [`trackr_network()`](https://hamishgibbs.github.io/rtrackr/reference/trackr_network.html). For more information, please see [getting started](https://hamishgibbs.github.io/rtrackr/articles/getting_started.html). Note that this record has two "root" records (it is the child of two starting points in the processing chain).

```{r}
target_id <- df$trackr_id[1]
trackr_lineage(target_id, trackr_dir)

lineage_fn <- paste0(trackr_dir, '/', target_id, '_lineage.json')

trackr_network(lineage_fn)
```

### Clean up
```{r results="hide"}
clean_trackr_dir(trackr_dir)
```

*Article by Hamish Gibbs `r Sys.time()`. To report a problem with this package, please [create an issue](https://github.com/hamishgibbs/rtrackr/issues) on [GitHub](https://github.com/hamishgibbs/rtrackr).*