---
title: "Single history workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Single history workflow}
  %\usepackage[UTF-8]{rtrackr}
---

```{r echo=F, results="hide"}
suppressPackageStartupMessages({
  library(kableExtra)
  library(rtrackr)
  library(networkD3)
})
```

### Setting up
Below is an example of a more complicated workflow, starting from a simple input file, to be used as an example of a full `rtrackr` workflow.

We will define and log a simple dataframe for this example. For more information, please see [Getting started](https://hamishgibbs.github.io/rtrackr/articles/getting_started.html).

```{r}
df <- data.frame(a = c('a', 'b', 'c'), b = c(1, 2, 3))
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
```{r}
trackr_dir <- '~/Documents/Personal/trackr_dev/trackr_dir'

#setup a new trackr
df <- trackr_new(df, trackr_dir = trackr_dir, timepoint_message = 'Start', log_data = T)

```

### Workflow

```{r}
#make some change to the data 
df <- df %>% dplyr::mutate(b = b + 1)

df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'Change point #1', log_data = T)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
```{r}

#repeat for subsequent processing steps
df <- df %>% dplyr::mutate(b = b + 300)

df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'Change point #2', log_data = T)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
```{r}
#splitting a single record into multiples
df <- rbind(df, df %>% dplyr::mutate(b = b -300))

df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'Dividing rows', log_data = T)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
```{r}
#summarising data
df <- df %>% dplyr::group_by(a)

df <- trackr_summarise(df, n = dplyr::n())

df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'Summarising rows', log_data = T)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
```{r}
df <- df %>% dplyr::mutate(n = n**runif(3))

df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'Further Processing', log_data = T)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
```{r}
df <- rbind(df, df %>% dplyr::mutate(n = n -60), df %>% dplyr::mutate(n = n * 2), df %>% dplyr::mutate(n = n * 56))

df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'Really messing with things', log_data = T)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
```{r}
df <- df %>% dplyr::mutate(n = n + 1)

df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'added 1', log_data = T)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
```{r}
df <- df %>% dplyr::group_by(a) %>% 
  trackr_summarise(n =  dplyr::n())

df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'grouped', log_data = T)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>
```{r}
df <- df %>% dplyr::mutate(n = n + 2000)

df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'added 2000', log_data = T)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>

### Results
Inspect the full processing chain of a single record with [`trackr_lineage()`](https://hamishgibbs.github.io/rtrackr/reference/trackr_lineage.html) and [`trackr_network()`](https://hamishgibbs.github.io/rtrackr/reference/trackr_network.html).

```{r}
target_id <- df$trackr_id[1]
trackr_lineage(target_id, trackr_dir)

lineage_fn <- paste0(trackr_dir, '/', target_id, '_lineage.json')

trackr_network(lineage_fn)
```

### Clean up
```{r results="hide"}
clean_trackr_dir(trackr_dir)
```

*Article by Hamish Gibbs `r Sys.time()`. To report a problem with this package, please create an [issue](https://github.com/hamishgibbs/rtrackr/issues).*