---
title: "Summarising data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Summarising data}
  %\usepackage[UTF-8]{rtrackr}
---


```{r echo=F, results="hide"}
suppressPackageStartupMessages({
  library(kableExtra)
  library(rtrackr)
  library(networkD3)
})
```

### Overview

`rtrackr` provides data logging for every record in a dataset throughout the processing chain. In most cases, when records are altered or one record is divided to multiple records, `rtrackr` will simply assign a new trackr id and log changes when a record is updated. 

When data is summarised, on the other hand (multiple records become a single record), `rtrackr` needs to record the `trackr_ids` of all parent records. 

[`trackr_summarise()`](https://hamishgibbs.github.io/rtrackr/reference/trackr_summarise.html) provides a convenient way to summarise data without losing information in the `trackr_id` column.

[`trackr_summarise()`](https://hamishgibbs.github.io/rtrackr/reference/trackr_summarise.html) works by combining all parent ids into one row, separated by a ", ". The same operation would work for combining records manually outside of R. 

### Example workflow

We will use a simple workflow To demonstrate the use of [`trackr_summarise()`](https://hamishgibbs.github.io/rtrackr/reference/trackr_summarise.html) in a data processing chain. Following on from [Getting started](https://hamishgibbs.github.io/rtrackr/articles/getting_started.html), we will create a new dataset, log a new processing timepoint with [`trackr_new()`](https://hamishgibbs.github.io/rtrackr/reference/trackr_new.html).

```{r}
trackr_dir <- '~/Documents/trackr_dir'
df <- data.frame(a = c('a', 'b', 'c'), b = c(1, 2, 3))
df <- trackr_new(df, trackr_dir = trackr_dir)
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>

Now, we will bind the dataset to itself, and make some changes to one version. 

```{r}
df <- rbind(df, df %>% dplyr::mutate(b = b + 1))
df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'Merged dataframes')
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>

`trackr_summarise` is a simple wrapper around [`dplyr::summarise`](https://dplyr.tidyverse.org/reference/summarise.html) and accepts the same arguments. 

```{r}
df <- df %>% 
  dplyr::group_by(a) %>% 
  trackr_summarise(n = dplyr::n())

```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>

Now, we can log a new timepoint with [`trackr_timepoint()`](https://hamishgibbs.github.io/rtrackr/reference/trackr_timepoint.html).

```{r}
df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'Summarised dataframes')
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>

We will make and log one more change, to better visualize the effect of the summarise operation.

```{r}
df <- df %>% dplyr::mutate(n = n + 100)
df <- trackr_timepoint(df, trackr_dir = trackr_dir, timepoint_message = 'Added 100')
```
<details><summary>Details</summary>
```{r, echo = F}
kable(df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
</details>

To visualize this operation on one record, we create a `trackr_lineage` and `trackr_network`. See [Getting started](https://hamishgibbs.github.io/rtrackr/articles/getting_started.html) for more information. 

```{r}
target_id <- df$trackr_id[1]
trackr_lineage(target_id, trackr_dir)

lineage_fn <- paste0(trackr_dir, '/', target_id, '_lineage.json')

trackr_network(lineage_fn)
```

### Clean up

```{r results="hide"}
clean_trackr_dir(trackr_dir)
```

*Article by Hamish Gibbs `r Sys.time()`. To report a problem with this package, please [create an issue](https://github.com/hamishgibbs/rtrackr/issues) on [GitHub](https://github.com/hamishgibbs/rtrackr).*